Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-09-13 intro.tex
words on factorisation and pdfs

\label{chapter:intro}
\section{The Standard Model of particle physics}
    The Standard Model of particle physics (SM) is our current
    best working theory to describe all known elementary particles
    as well as three of the four fundamental forces. Developed
    predominantly in the latter half of the 20th century, it is
    one of the most well tested theories we have in science today.
    Some highlights include the highly precise predictions of the
    anomalous magnetic moment of the electron, agreeing with experimental
    measurements to more than 10 significant figures \cite{Aoyama:2017uqe}.
    and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
    and CMS experiments \cite{CMS:2012qbp} at the Large Hadron Collider
    (LHC) which was theorised decades prior.

    The SM is a gauge quantum field theory (QFT) where particles
    are described as excitations in quantum fields. It can be specified
    by the gauge group
    \begin{equation}\label{eqn:SM_gauge}
        \mathrm{SU}(3)_{C} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}\, ,
    \end{equation}
    where subscripts denote the charges of the gauge groups. The first gauge
    group with colour charge $C$ describes the interactions of the
    strong force within the theory of quantum chronodynamics (QCD), which we
    will elaborate more on in the next section. The second
    and third gauge groups represent the electroweak sector of the SM with
    charges $L$ for left and $Y$ for hypercharge. Under electroweak
    spontaneous symmetry breaking (EWSB), this product becomes
    \begin{equation}\label{eqn:SM_SSB}
        \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y} \xrightarrow{\mathrm{EWSB}} \mathrm{U}(1)_{\mathrm{EM}} \, ,
    \end{equation}
    giving rise to the electromagnetic and weak forces we observe.
    The effect of gravity is considered to be negligible on the scales
    considered in the SM and so it is not described.
    Each of the three fundamental forces described by the SM is mediated
    by the exchange of a gauge boson. The gauge bosons mediating the strong
    and EM force are the massless gluon $g$ and massless photon $\gamma$,
    respectively. For the weak force the gauge bosons are the $W^{\pm}$
    and $Z^{0}$ bosons which attain a mass through the Higgs mechanism \cite{Englert:1964et,Higgs:1964pj,Guralnik:1964eu}
    during EWSB, which elucidates the Higgs boson $H$.

@@

\section{Introduction to quantum chromodynamics}
    Quantum chronodynamics (QCD) is the sector of the SM that
    describes the strong interaction. 
    QCD is a non-abelian gauge theory with gauge group
    $\mathrm{SU}(3)_{C}$ where the charge is named colour.
    The gauge and fermion part of the QCD Lagrangian is
    \begin{equation}\label{eqn:L_QCD}
        \mathcal{L}_{\mathrm{QCD}} = -\dfrac{1}{4}F^{a}_{\mu\nu}F^{a, \, \mu\nu}
        + \sum_{f} \bar{\psi}_{i}^{f}(\mathrm{i}\slashed{D}_{ij} - \delta_{ij}m_{f})\psi_{j}^{f} \, ,
    \end{equation}
    where twice repeated indices are summed over.
    The fields $\psi_{i}^{f}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
    representing quarks (antiquarks) with flavours $f$: 
    up $u$, down $d$, strange $s$, charm $c$, top $t$, and bottom $b$,
    with masses $m_{f}$. They transform under the fundamental
    representation with indices $i, j \in \{1, 2, 3\}$, named colour indices.
    The gauge fields $A^{a}_{\mu}$, corresponding to gluons, 
    appear in the quark covariant derivative
    \begin{equation}\label{eqn:covariant_deriv}
        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{\mathrm{s}}T^{a}_{ij}A^{a}_{\mu} \, ,
    \end{equation}
    and the gauge field strength tensor
    \begin{equation}\label{eqn:F_munu}
        F^{a}_{\mu\nu} = \partial_{\mu}A^{a}_{\nu} - \partial_{\nu}A^{a}_{\mu} + g_{\mathrm{s}} f^{abc}A^{b}_{\mu}A^{c}_{\nu} \, .
    \end{equation}
    Gauge fields transform under the adjoint representation, meaning
    the adjoint indices $a,b,c \in \{1,...,8\}$. $T^{a}_{ij}$ are the
    group generators in a fundamental representation. In SU(3) it is
    common to write the group generators as
    \begin{equation}\label{eqn:group_generators}
        T^{a}_{ij} = \dfrac{1}{2}\lambda^{a}_{ij}
    \end{equation}
    where $\lambda^{a}_{ij}$ are the Gell-Mann matrices \cite{Gell-Mann:1962yej}.
    The generators of the group have to obey the Lie algebra
    \begin{equation}\label{eqn:lie_algebra}
        [T^{a}, T^{b}] = \mathrm{i}f^{abc}T^{c} \, ,
    \end{equation}
    where $f^{abc}$ are the structure constants of SU(3).
    The gauge coupling of the group $g_{\mathrm{s}}$
    is a dimensional free parameter of the theory, usually
    written as
    \begin{equation}\label{eqn:alpha_s}
        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, ,
    \end{equation}
    for reasons which will become clear when we discuss
    scattering amplitudes.

@@

    Now that we have written down the QCD Lagrangian,
    we can examine the terms to write down how the
    gauge bosons interact with the fermions. A convenient
    way to do this is with Feynman rules, which can be
    pieced together to form Feynman diagrams. We will see
    that Feynman diagrams are a simple, but powerful, tool
    to systematically build all the possible ways a process
    can occur. This will be discussed further in the
    scattering amplitude section.

@@

\section{Scattering amplitudes and cross-sections}
    In collider experiments we typically collide
    two beams, consisting of bundles of energetic
    particles, and analyse the resulting products
    of these collisions. To model collisions in
    a collider experiment, we look at the specific
    case of two particles colliding into $n$ particles.
    To mathematically describe these
    scattering processes, scattering amplitudes are used.
    For an initial-state
    $|i\rangle$ and final-state $|f\rangle$
    the scattering amplitude can be written as
    \begin{equation}\label{eqn:scattering}
        \langle f | S | i \rangle \, ,
    \end{equation}
    where the scattering matrix, or $S$-matrix
    can be decomposed into an identity matrix
    and a transfer matrix $\mathcal{T}$,
    \begin{equation}\label{eqn:S_matrix}
        S = \mathbbm{1} + \mathrm{i}\mathcal{T} \, .
    \end{equation}
    The $S$-matrix encodes all the information
    about how the initial-state will evolve
    into the final-state over time. By writing
    the $S$-matrix in this way, all the interactions
    are separated into the transfer matrix $\mathcal{T}$,
    as the identity matrix describes the free theory.
    By imposing a momentum conserving constraint
    on the $S$-matrix, the matrix element,
    $\mathcal{M}$, is defined in the following expression
    \begin{equation}\label{eqn:matrix_element}
        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, .
    \end{equation}
    where we take $p_{a}$ and $p_{b}$ to be
    the momenta of the two colliding
    particles in the initial-state, and $p_{f}$ to be the
    momenta of the $n$-particle final states.
    The transition probability of this scattering process
    is the modulus squared of the scattering amplitude
    \begin{equation}\label{eqn:S_prob}
        P = |\langle f | S | i \rangle|^{2} \propto |\langle f | \mathcal{M} | i \rangle 
    \end{equation}
    where $|\langle f | \mathcal{M} | i \rangle|^{2}  \equiv |\mathcal{M}|^{2}$
    is the matrix element squared.

@@

    Since detectors at collider experiments
    have limited resolution, single scattering
    events are not measured. Instead we consider
    the cross-section, $\sigma$, which is a property
    of the particles being scattered, and is independant
    of the way the experiment is carried out.
    The cross-section directly relates the
    matrix elements which we calculate in our QFT,
    to measurable observables at experiments.
    In practice, it is more useful to consider the
    differential cross-section, $\mathrm{d}\sigma$
    where the cross-section can be differential
    in quantities such as energies and angles.
    This is because the final-state particles
    will be produced with a range of, for example,
    energies and momenta so the cross-section would
    be vanishing for an exact momentum configuration.

@@

    $\mathrm{d}\Phi_{n}$ is the Lorentz-invariant
    phase-space which contains all the possible
    configurations of the $n$-particle final state.
    Absorbing the momentum conserving $\delta$-function
    from (\ref{eqn:matrix_element}),
    we can write it as
    % \begin{equation}\label{eqn:dlips}
    %     \mathrm{d}\Phi_{n} = (2\pi)^{d}\delta^{(d)}\left(p_{a} + p_{b} - \sum_{f} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{d}p_{f}}{(2\pi)^{d-1}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
    % \end{equation}
    \begin{equation}\label{eqn:dlips_4d}
        \mathrm{d}\Phi_{n} = (2\pi)^{4}\delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{4}p_{f}}{(2\pi)^{3}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
    \end{equation}
    where the second $\delta$-function restricts
    the final state particles to be on mass-shell
    (on-shell). The step function selects only
    the positive energy solution for this constraint.
    \begin{itemize}
        \item Matrix elements are calculated at fixed order in perturbative QFT -> Need to mentioned what perturbative even is...
        \item How can QCD be perturbative when $\alpha_{s}$ is large? running, beta functions, asymptotic freedom
        \item Why don't we see quarks and gluons at colliders? Hard/soft factorisation, pdfs, DGLAP, confinement, infrared safe observables
        \item Relate hadronic cross-section to partonic cross-section.
    \end{itemize}

\section{Experiments and Phenomonology}
\subsection{Observables}
\begin{itemize}
    \item Experiments do not measure single particles,
    they measure hadrons which is far from the collision
    of protons.
    \item Factorisation of hard and soft scales leads
    to the concepts of PDFs (DGLAP equation), partonic
    cross-sections (Feynman diagrams), parton showers,
    and hadronisation.
    \item Link back to matrix elements and how they
    are integrated to give you cross-sections which
    can be compared with experiments.
    \item Probably need to talk about infrared safe
    observables and jets.
\end{itemize}
\section{Monte Carlo event generator overview}
\subsection{Event unweighting}
\begin{itemize}
    \item MC generators are the main tools that
    people use to compare with experimental data.
    There are many groups that have their own tool.
    \item The usage of MC tools came about because
    the multi-dimensional phase-space integrals
    are too complex to do analytically.
    \item This doesn't mean that they only do
    the integrals, the event generators basically
    contain all the tools from phase-space generation
    to matrix element evaluation, to parton showers
    to hadronisation.
    \item Have a look at Katherina thesis to see
    how much I can skip because only really want to
    introduce unweighting. Probably along the lines
    of event weight is statistically sampled vs
    what they measure in real life. Therefore we need
    to unweight events.
\end{itemize}
\end{document}

>>>>

\section{The Standard Model of particle physics}
    The Standard Model of particle physics (SM) is our current
    best working theory to describe all known elementary particles
    as well as three of the four fundamental forces. Developed
    predominantly in the latter half of the 20th century, it is
    one of the most well tested theories we have in science today.
    Some highlights include the highly precise predictions of the
    anomalous magnetic moment of the electron, agreeing with experimental
    measurements to more than 10 significant figures \cite{Aoyama:2017uqe}.
    and the discovery of the Higgs boson in 2012 by the ATLAS \cite{ATLAS:2012yve}
    and CMS experiments \cite{CMS:2012qbp} at the Large Hadron Collider
    (LHC) which was theorised decades prior.

    The SM is a gauge quantum field theory (QFT) where particles
    are described as excitations in quantum fields. It can be specified
    by the gauge group
    \begin{equation}\label{eqn:SM_gauge}
        \mathrm{SU}(3)_{c} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}\, ,
    \end{equation}
    where subscripts denote the charges of the gauge groups. The first gauge
    group with colour charge $C$ describes the interactions of the
    strong force within the theory of quantum chronodynamics (QCD), which we
    will elaborate more on in the next section. The second
    and third gauge groups represent the electroweak sector of the SM with
    charges $L$ for left and $Y$ for hypercharge. Under electroweak
    spontaneous symmetry breaking (EWSB), this product becomes
    \begin{equation}\label{eqn:SM_SSB}
        \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y} \xrightarrow{\mathrm{EWSB}} \mathrm{U}(1)_{\mathrm{EM}} \, ,
    \end{equation}
    giving rise to the electromagnetic and weak forces we observe.
    The effect of gravity is considered to be negligible on the scales
    considered in the SM and so it is not described.
    Each of the three fundamental forces described by the SM is mediated
    by the exchange of a gauge boson. The gauge bosons mediating the strong
    and EM force are the massless gluon $g$ and massless photon $\gamma$,
    respectively. For the weak force the gauge bosons are the $W^{\pm}$
    and $Z^{0}$ bosons which attain a mass through the Higgs mechanism \cite{Englert:1964et,Higgs:1964pj,Guralnik:1964eu}
    during EWSB, which elucidates the Higgs boson $H$.

@@

\section{Introduction to quantum chromodynamics}
    \subsection{The QCD Lagrangian}
    Quantum chronodynamics (QCD) is the sector of the SM that
    describes the strong interaction. 
    QCD is a non-Abelian gauge theory with gauge group
    $\mathrm{SU}(3)_{c}$ where the charge is named colour.
    The gauge and fermion part of the QCD Lagrangian is
    \begin{equation}\label{eqn:L_QCD}
        \mathcal{L}_{\mathrm{QCD}} = -\dfrac{1}{4}F^{a}_{\mu\nu}F^{a, \, \mu\nu}
        + \sum_{f} \bar{\psi}_{i}^{f}(\mathrm{i}\slashed{D}_{ij} - \delta_{ij}m_{f})\psi_{j}^{f} \, ,
    \end{equation}
    where twice repeated indices are summed over.
    The fields $\psi_{i}^{f}$ ($\bar{\psi}_{i}^{f}$) are the fermions (antifermions),
    representing quarks (antiquarks) with flavours $f$: 
    up $u$, down $d$, strange $s$, charm $c$, top $t$, and bottom $b$,
    with masses $m_{f}$. They transform under the fundamental
    representation with indices $i, j \in \{1, 2, 3\}$, named colour indices.
    The gauge fields $A^{a}_{\mu}$, corresponding to gluons, 
    appear in the quark covariant derivative
    \begin{equation}\label{eqn:covariant_deriv}
        (D_{\mu})_{ij} = \delta_{ij}\partial_{\mu} - ig_{\mathrm{s}}T^{a}_{ij}A^{a}_{\mu} \, ,
    \end{equation}
    and the gauge field strength tensor
    \begin{equation}\label{eqn:F_munu}
        F^{a}_{\mu\nu} = \partial_{\mu}A^{a}_{\nu} - \partial_{\nu}A^{a}_{\mu} + g_{\mathrm{s}} f^{abc}A^{b}_{\mu}A^{c}_{\nu} \, .
    \end{equation}
    Gauge fields transform under the adjoint representation, 
    which is an $N_{c}^{2}-1$ dimensional representation, meaning
    the adjoint indices $a,b,c \in \{1,...,8\}$. $T^{a}_{ij}$ are the
    group generators in a fundamental representation. In SU(3) it is
    common to write the group generators as
    \begin{equation}\label{eqn:group_generators}
        T^{a}_{ij} = \dfrac{1}{2}\lambda^{a}_{ij}
    \end{equation}
    where $\lambda^{a}_{ij}$ are the Gell-Mann matrices \cite{Gell-Mann:1962yej}.
    The generators of the group have to obey the Lie algebra
    \begin{equation}\label{eqn:lie_algebra}
        [T^{a}, T^{b}] = \mathrm{i}f^{abc}T^{c} \, ,
    \end{equation}
    where $f^{abc}$ are the structure constants of SU(3).
    The gauge coupling of the group $g_{\mathrm{s}}$
    is a dimensionless free parameter of the theory.
    It is common to use the strong coupling constant
    instead, 
    \begin{equation}\label{eqn:alpha_s}
        \alpha_{\mathrm{s}} = \dfrac{g_{\mathrm{s}}^{2}}{4\pi} \, .
    \end{equation}
    The strong coupling constant, is in fact not constant, but
    is dependent on the energy scale of the process, $\mu$. This
    is due to the process of renormalisation which will be discussed
    in section ...

@@

    \subsection{QCD Feynman rules}
    Now that we have written down the QCD Lagrangian,
    we can examine the terms to write down how the
    gauge bosons interact with the fermions. A convenient
    way to do this is with Feynman rules, which can be
    pieced together to form Feynman diagrams. We will see
    that Feynman diagrams are a simple, but powerful, tool
    to systematically build all the possible ways a process
    can occur. This will be discussed further in the
    scattering amplitude section.

@@

\section{Factorisation theorem}
    At the LHC, collisions occur between two protons,
    which have constituent quarks and gluons, collectively
    named partons.
    The application of QCD to describe phenomenon
    in these proton-proton collisions rests on the use of
    the factorisation theorem which enables the
    separation of soft and hard energy scales.
    Within the proton, there are composite quarks
    and gluons that are constantly absorbed and
    emitted on a timescale inversely proportional
    to the mass of the proton (or hadron in general).
    This fluctuation timescale is much longer
    than the timescale of the interaction between
    an incoming probing parton from another proton in a
    highly energetic collision, like at the LHC.
    During the collision, or the hard scattering,
    the probe is able to interact with a
    parton that is effectively frozen.
    The probing parton knows nothing of the
    proton being probed, except for the fact
    that it collided with a parton carrying
    a fraction of the proton momentum. The
    distribution of partons within a parton
    is process independent and a fundamental
    property of the proton, meaning it can be
    separated from the actual scattering between
    the partons.

    This heuristic argument motivates the form of
    the factorisation equation, where the production of
    particles from a hadronic collision, the hadronic
    cross-section, can be
    written as
    \begin{equation}\label{eqn:hadronic_cs}
        \sigma_{AB \rightarrow n} = \sum_{a, b}\int_{0}^{1} \mathrm{d}x_{a}\mathrm{d}x_{b} \; f_{a/A}(x_{a},\mu_{\mathrm{F}})f_{b/B}(x_{b}, \mu_{\mathrm{F}}) \; \hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}}) + \mathcal{O}\left(\dfrac{\Lambda_{\mathrm{QCD}}}{Q}\right)\, ,
    \end{equation}
    where $a$ is the parton identified from hadron $A$,
    and $b$ is the parton identified from hadron $B$.
    This factorisation is not exact as indicated by the
    correction term inversely proportional to characteristic
    hard scale $Q$. However, for high energy collisions, where
    $Q \gg \Lambda_{\mathrm{QCD}}$, this term and higher order
    terms are suppresed.
    In fact, the factorisation equation has only been
    proven for a few specific cases (DIS etc, cite here).

    $f_{a/h}(x, \mu)$ are parton distribution functions
    (PDFs) which depend on the momenta fractions $x$ parton $a$
    has with respect to its parent hadron $h$, at a scale $\mu$,
    usually taken to be the factorisation scale $\mu_{\mathrm{F}}$,
    the interface between soft and hard physics.
    The interpretation of PDFs at leading order are
    as probability distributions. $f_{a/h}(x, \mu)$ is
    the probability of finding parton $a$ within $h$
    carrying a momentum fraction $x$ at the energy scale $\mu$.
    PDFs are non-perturbative objects that encapsulate
    the soft effects in a scattering process below energy
    $\mu_{\mathrm{F}}$. They are non-perturbative because
    they are determined by fitting to experimental data,
    instead of calculated in perturbation theory.
    Multiple groups release PDF fits (cite groups here).
    In practice, these PDF fits
    are accessed through the LHAPDF interface (cite here).
    The evolution of PDFs between factorisation scales is possible
    through the use of the Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP)
     equations (cite here).

    $\hat{\sigma}_{ab \rightarrow n}(Q, \mu_{\mathrm{F}}, \mu_{\mathrm{R}})$
    is the partonic cross-section which descibes the interaction of
    partons $a$ and $b$ scattering into $n$ particles,
    with the scattering occurs at energy scale $Q$.
    Notice that the partonic cross-section depends on
    both the factorisation scale, $\mu_{\mathrm{F}}$ and
    the renormalisation scale, $\mu_{\mathrm{R}}$.
    {\color{red} Need to mention how these scales
    are unphysical but there is a residual dependence
    on them at fixed-order calculations. Scale
    variations are a conventional way to determine
    the effects of higher order terms.
    Renormalisation will be discussed in Chapter 2
    when divergences are introduced.}

    The discussion of how $\hat{\sigma}_{ab \rightarrow n}$
    is computed is discussed in detail in the next section.

\section{Scattering amplitudes and cross-sections}
    {\color{red} Cross-sections here are all
    partonic cross-sections. Need to talk about
    how they can be calculated order by order
    in perturbation theory with Feynman diagrams.}

    In collider experiments we typically collide
    two beams, consisting of bundles of energetic
    particles, and analyse the resulting products
    of these collisions. To model collisions in
    a collider experiment, we look at the specific
    case of two particles colliding into $n$ particles.
    To mathematically describe these
    scattering processes, scattering amplitudes are used.
    For an initial-state
    $|i\rangle$ and final-state $|f\rangle$
    the scattering amplitude can be written as
    \begin{equation}\label{eqn:scattering}
        \langle f | S | i \rangle \, ,
    \end{equation}
    where the scattering matrix, or $S$-matrix
    can be decomposed into an identity matrix
    and a transfer matrix $\mathcal{T}$,
    \begin{equation}\label{eqn:S_matrix}
        S = \mathbbm{1} + \mathrm{i}\mathcal{T} \, .
    \end{equation}
    The $S$-matrix encodes all the information
    about how the initial-state will evolve
    into the final-state over time. By writing
    the $S$-matrix in this way, all the interactions
    are separated into the transfer matrix $\mathcal{T}$,
    as the identity matrix describes the free theory.
    By imposing a momentum conserving constraint
    on the $S$-matrix, the matrix element,
    $\mathcal{M}$, is defined in the following expression
    \begin{equation}\label{eqn:matrix_element}
        \langle f | \mathcal{T} | i \rangle = (2\pi)^{4} \delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \mathcal{M} \, ,
    \end{equation}
    where we take $p_{a}$ and $p_{b}$ to be
    the momenta of the two colliding
    particles in the initial-state, and $p_{f}$ to be the
    momenta of the $n$-particle final states.
    The transition probability of this scattering process
    is the modulus squared of the scattering amplitude
    \begin{equation}\label{eqn:S_prob}
        P = |\langle f | S | i \rangle|^{2} \propto |\langle f | \mathcal{M} | i \rangle 
    \end{equation}
    where $|\langle f | \mathcal{M} | i \rangle|^{2}  \equiv |\mathcal{M}|^{2}$
    is the matrix element squared.

@@

    Since detectors at collider experiments
    have limited resolution, single scattering
    events are not measured. Instead we consider
    the cross-section, $\sigma$, which is a property
    of the particles being scattered, and is independent
    of the way the experiment is carried out.
    The cross-section directly relates the
    matrix elements which we calculate in our QFT,
    to measurable observables at experiments.
    In practice, it is more useful to consider the
    differential cross-section, $\mathrm{d}\sigma$,
    where the cross-section can be differential
    in quantities such as energies and angles.
    This is because the final-state particles
    will be produced with a range of, for example,
    energies and momenta so the cross-section would
    be vanishing for an exact momentum configuration.

@@

    $\mathrm{d}\Phi_{n}$ is the Lorentz-invariant
    phase-space which contains all the possible
    configurations of the $n$-particle final state.
    Absorbing the momentum conserving $\delta$-function
    from (\ref{eqn:matrix_element}),
    we can write it as
    % \begin{equation}\label{eqn:dlips}
    %     \mathrm{d}\Phi_{n} = (2\pi)^{d}\delta^{(d)}\left(p_{a} + p_{b} - \sum_{f} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{d}p_{f}}{(2\pi)^{d-1}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
    % \end{equation}
    \begin{equation}\label{eqn:dlips_4d}
        \mathrm{d}\Phi_{n} = (2\pi)^{4}\delta^{4}\left(p_{a} + p_{b} - \sum_{f=1}^{n} p_{f}^{\mu}\right) \prod_{f=1}^{n} \dfrac{\mathrm{d}^{4}p_{f}}{(2\pi)^{3}}\delta(p_{f}^{2} - m_{f}^{2})\theta(E_{f}) \, ,
    \end{equation}
    where the second $\delta$-function restricts
    the final state particles to be on mass-shell
    (on-shell). The step function selects only
    the positive energy solution for this constraint.

\section{Monte Carlo event generator overview}
\subsection{Event unweighting}
\begin{itemize}
    \item MC generators are the main tools that
    people use to compare with experimental data.
    There are many groups that have their own tool.
    \item The usage of MC tools came about because
    the multi-dimensional phase-space integrals
    are too complex to do analytically.
    \item This doesn't mean that they only do
    the integrals, the event generators basically
    contain all the tools from phase-space generation
    to matrix element evaluation, to parton showers
    to hadronisation.
    \item Have a look at Katherina thesis to see
    how much I can skip because only really want to
    introduce unweighting. Probably along the lines
    of event weight is statistically sampled vs
    what they measure in real life. Therefore we need
    to unweight events.
\end{itemize}
\end{document}
